
# Reproducible Research: Peer Assessment 1

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Loading and preprocessing the data

Loading the data and transforming the data into a format suitable for analysis:

```{r}
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata/data/activity.zip"
download.file( fileUrl,destfile="activity.zip" )
system( "unzip -o activity.zip" )
activity <- read.csv("activity.csv")
```
```{r,results="hide"}
library( dplyr )
activity <- tbl_df( activity )
```
```{r}
```


### Mean total number of steps taken per day

I ignore the missing values in the dataset for this part of the assignment.

1. Calculating the total number of steps taken per day: 
```{r}
plot_data <- group_by( activity, date ) %>%
    summarize( steps_sum = sum(steps,na.rm=TRUE) )
```
2. Making a histogram of the total number of steps taken each day:
```{r}
#png( file="plot_sum_date.png",width = 480,height = 480 )
with(plot_data, barplot( plot_data$steps_sum, names.arg=plot_data$date, xlab="Date", ylab="Steps", main="Total steps each day" ))
#dev.off()
```
3. Calculating and reporting the mean and median of the total number of steps taken per day:
```{r}
steps_mean <- mean( plot_data$steps_sum,na.rm=TRUE )
print( paste("Mean of total number of steps taken per day:",steps_mean ) )
steps_median <- median( plot_data$steps_sum,na.rm=TRUE )
print( paste("Median of total number of steps taken per day:",steps_median ) )
```

### Average daily activity pattern

1. Making a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis):
```{r}
plot_data <- group_by( activity, interval ) %>%
    summarize( steps_sum    =    mean(steps, na.rm=TRUE))
#png( file="plot_sum_time.png",width = 480,height = 480 )
with(plot_data, plot( interval, steps_sum, type="l", xlab="5-min interval", ylab="Steps", main="Total steps each interval" ))
#dev.off()
```
2. Getting 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps:
```{r}
max_int <- plot_data[ which( plot_data$steps_sum == max(plot_data$steps_sum)), ]
print( paste("5-minute interval contains the maximum number of steps is",max_int$interval) )
```

### Imputing missing values

This dataset contains a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

1. Calculating and reporting the total number of missing values in the dataset:
```{r}
print(paste("Total number of missing values in the dataset is",sum(is.na(activity$steps))))
```
2. Replacing all of the missing values in the dataset with the mean for that 5-minute interval:
```{r}
activity_f <- group_by( activity, interval )
activity_f <- mutate( activity_f, steps_m = mean(steps,na.rm=TRUE))
```
3. Creating a new dataset that is equal to the original dataset but with the missing data filled in:
```{r}
activity_f <- activity_f %>%
    mutate( steps = if_else(is.na(steps), as.integer(steps_m), steps) )
activity_f <- select(activity_f, -steps_m)
```
4. Making a histogram of the total number of steps taken each day and calculating and reporting the mean and median total number of steps taken per day:
```{r}
plot_data <- group_by( activity_f, date ) %>%
    summarize( steps_sum    =    sum(steps))
#png( file="plot_sum_date_na.png",width = 480,height = 480 )
with(plot_data, barplot( plot_data$steps_sum, names.arg=plot_data$date, xlab="Date", ylab="Steps", main="Total steps each day after imputing missing values" ))
#dev.off()
```
```{r}
steps_mean <- mean( plot_data$steps_sum,na.rm=TRUE )
print( paste("Mean of total number of steps taken per day after imputing missing values:",steps_mean ) )
steps_median <- median( plot_data$steps_sum,na.rm=TRUE )
print( paste("Median of total number of steps taken per day after imputing missing values:",steps_median ) )
```
